---
title: "Forcasting Oil Price"
author: "Mohammed AlKhamis"
date: "11/9/2021"
output:
  html_document: default
  pdf_document: default
---

# Importing Needed Packages 
```{r setup, include=FALSE}
library(TSstudio)
library(tidyverse)
library(forecast)
library(dygraphs)
library(ggfortify)
library(fpp2)


```

# Importing the Data 

```{r}
price <- read_csv("data/oil_price.csv")
head(price)
# Convert the data frame to time series 

price_ts <- price %>% dplyr::select(price = cost) %>% ts( start = c(1946, 1), frequency = 12)

```

# Exploring 

```{r}
dygraph(price_ts, 
        main  = "World Monthly Oil price ",
        ylab  = "price  in $",
        xlab  = "Months") %>% 
  dyRangeSelector()
```

```{r}
ts_decompose(price_ts) # illustrating seasonality as well as other parameters 
```

```{r}
ts_cor(price_ts) # Illustrating the lag seasonality 
acf(price_ts)
```

```{r}
ts_seasonal(price_ts, type = "normal")

```
# Time Series Models

## Seasonal Naive model
```{r}
dy <- diff(price_ts) # taking the first difference of the time series to eliminate the trend and make it stationary 
autoplot(dy)



fit<- snaive(dy) #Residual SD: 6.9603 
print(summary(fit))
checkresiduals(fit)
```

## ETS Model
```{r}
fit_ets<- ets(price_ts) #Residual SD: 0.089
print(summary(fit_ets))
checkresiduals(fit_ets)
```

## ARIMA Model 
```{r}
fit_arima<- auto.arima(price_ts, d=1) #Residual SD: 4.682948 
print(summary(fit_arima))
checkresiduals(fit_arima)

```

# Forcasting 
I choosed the ETS since it has the lowest residual

```{r}

fcs <- forecast(fit_ets, h= 12*10-11)# forcasting 10 years 
plot_forecast(fcs)
```
The prediction shows an increase in the oil price for the next 10 years. Which make sense since the 
oil is still going to be heavily used by lots of countries. Also, the oil use is going to change as well from 
electricity production to .....