---
title: "Forcasting Oil Consumption"
author: "Mohammed AlKhamis"
date: "11/9/2021"
output: html_document
---
# Importing Needed Packages 
```{r setup, include=FALSE}
library(TSstudio)
library(tidyverse)
library(forecast)
library(dygraphs)
library(tidyverse)
library(readxl)
library(tidyquant)
library(mice)
library(visdat)
library(naniar)
library(plotly)
library(ggplot2)
library(tidymodels)
library(ggfortify)

```

# Importing the Data 
```{r}
oil_consumption1 <- read_excel("data/oil-consumption-per-TBD.xlsx")

# change the class of the columns to numeric 
vec<- seq(2,57,1)
oil_consumption1[ , vec] <- apply(oil_consumption1[ , vec,drop=F], 2,           
                                 function(x) as.numeric(as.character(x)))

# Tidy my data 
oil_consumption1 <- oil_consumption1 %>% 
  pivot_longer(-c(country, per_region), names_to = "year", values_to = "oil_comsumption_in_EJ")

# Fill the missing values by back filling method  
oil_consumption1 <- oil_consumption1 %>% fill(oil_comsumption_in_EJ, .direction = "up")
sum(is.na(oil_consumption1))
```

```{r}
# head(oil_consumption1)
# Convert the data frame to time series 
consumption_ts <- oil_consumption1 %>% group_by(year) %>% summarise(Consumption_in_MBD = sum(oil_comsumption_in_EJ)/1000) %>% 
  dplyr::select(Consumption_in_MBD) %>%
  ts(start = 1965, frequency = 1)
```

# Exploring 

```{r}
dygraph(consumption_ts, 
        main  = "World Oil Consumption",
        ylab  = "Consumption in Million Barrels Daily",
        xlab  = "Years") %>% 
  dyRangeSelector()
```

```{r}
ts_cor(consumption_ts)# Illustrating the lag seasonality 

```
# Time Series Models

```{r}
dyy<- diff(consumption_ts) # taking the first difference of the time series to eliminate the trend and make it stationary
autoplot(dyy)
```

## Seasonal Naive model

```{r}
fit<- snaive(dyy) #Residual SD: 1.9801   
print(summary(fit))
checkresiduals(fit)
```

## ETS Model

```{r}

fit_ets<- ets(consumption_ts) #Residual SD: 0.0258
print(summary(fit_ets))
checkresiduals(fit_ets)
```

## ARIMA Model 
```{r}
fit_arima<- auto.arima(consumption_ts, d=1) #Residual SD: 1.843909
print(summary(fit_arima))
checkresiduals(fit_arima)
```
#AR model 
```{r}

library(vars)
library(mFilter)
library(tseries)
library(TSstudio)
library(forecast)
library(tidyverse)
VARselect(consumption_ts,lag.max = 10, type = "const")

AR <- ar(diff(consumption_ts), p= 2, type = "const") # Residual SD = 1.916507
print(summary(AR)) 
checkresiduals(AR)
```

# Forcasting Oil Consumtion 

I have chosen the ETS since it has the lowest residual
```{r}

fcs_cons <- forecast(fit_ets, h= 10) # forecasting 10 years 
autoplot(fcs_cons)
print(summary(fcs_cons))
plot_forecast(fcs_cons)
```
As you can see because of the pandemic the forecasting was miss leaded. Therefore, I tried to approach this problem 
from different side. I eliminated the 2020 drop (from the pandemic) and the forecast, lets see what happened



```{r}
oil_consumption2 <- read_excel("data/oil-consumption-per-TBD_2.xlsx")
 # change the class of the columns to numeric 
vec<- seq(2,56,1)
oil_consumption2[ , vec] <- apply(oil_consumption2[ , vec,drop=F], 2,           
                                 function(x) as.numeric(as.character(x)))

# Tidy my data 
oil_consumption2 <- oil_consumption2 %>% 
  pivot_longer(-c(country, per_region), names_to = "year", values_to = "oil_comsumption_in_MBD")

# therefore we will do a back fill 
oil_consumption2 <- oil_consumption2 %>% fill(oil_comsumption_in_MBD, .direction = "up")
sum(is.na(oil_consumption1))

```


```{r}
consumption_ts1 <- oil_consumption2 %>% group_by(year) %>% summarise(Consumption_in_MBD = sum(oil_comsumption_in_MBD)/1000) %>% 
 dplyr:: select(Consumption_in_MBD ) %>%
ts(, start = c(1965), frequency = 1)
plot(consumption_ts1)
```

## Seasonal Naive model

```{r}
fit1<- snaive(diff(consumption_ts1)) #Residual SD: 1.5193     
print(summary(fit1))
checkresiduals(fit1)
```

ETS Model 
```{r}
fit_ets1<- ets(consumption_ts1) #Residual SD: 1.3983
print(summary(fit_ets1))
checkresiduals(fit_ets1)
```


## ARIMA Model 
```{r}
fit_arima1<- auto.arima(consumption_ts1, d=1) #Residual SD: 1.246595
print(summary(fit_arima1))
checkresiduals(fit_arima1)
```
#AR model 
```{r}

library(vars)
library(mFilter)
library(tseries)
library(TSstudio)
library(forecast)
library(tidyverse)
VARselect(consumption_ts1,lag.max = 10, type = "const")

AR1 <- ar(diff(consumption_ts1), p= 2, type = "const") # Residual SD = 1.2907
AR1 <- ar(diff(consumption_ts1), p= 6, type = "const") # Residual SD = 1.2907

print(summary(AR)) 
checkresiduals(AR)
```

```{r}
# forecast
fcs_cons1 <- forecast(fit_arima1, h= 10)
autoplot(fcs_cons1)
print(summary(fcs_cons1))
plot_forecast(fcs_cons1)
```


```{r}
forcast_con1 <- fortify(fcs_cons1) 
forcast_con1[1:56, 3]

consumption_df1 <- data.frame(date  = forcast_con1$Index, consumption = c(forcast_con1 [ 1:56, 3], forcast_con1 [ 57:nrow(forcast_con1), 4]))
model <- consumption_df1

```

Now lets add the drop in 2020

```{r echo=FALSE}
  
#adjust
model[56, 2] <- 88.477
model[62:65, 2]  <- model[62:65, 2] -  seq(1,2,0.33)
plot(model, type= "l")
  

```






